2.3 Prob Model for PCR Positivity Estimation

### 2024-06-13 2_scripts/2.1 APL Data Quality Check_new.qmd
Update codes using APL v2 (unfinished) Further investigation on vaccination time gap

summary(RFD4682_v2$COLLECTION_DATE[!is.na(RFD4682_v2$`N-IgG_INTERP`)])
```

@@

```{r}
# Add encrypted IDe to vax and NAAT data
RFD4682_v2_vaxe<-merge(RFD4682_v2_vax, select(ULI_crosswalk, ULI_ID, ULI_IDe),
      by.x="ULI", by.y="ULI_ID", all.x = T, all.y = F)
# any missing?
table(is.na(RFD4682_v2_vaxe$ULI_IDe))

>>>>

table(RFD4682_v2$`N-IgG_INTERP`, useNA = 'ifany')
summary(RFD4682_v2$COLLECTION_DATE[!is.na(RFD4682_v2$`N-IgG_INTERP`)])
```

@@

## Unfinished ##
```{r}
# Add encrypted IDe to vax and NAAT data
RFD4682_v2_vaxe<-merge(RFD4682_v2_vax, select(ULI_crosswalk, ULI_ID, ULI_IDe),
      by.x="ULI", by.y="ULI_ID", all.x = T, all.y = F)
# any missing?
table(is.na(RFD4682_v2_vaxe$ULI_IDe))

### 2024-06-13 2_scripts/2.3 Prob Model for PCR pos.qmd
Update codes using APL v2 (unfinished) Further investigation on vaccination time gap

# The APL data
load(file = "../1_data/private/RFD4682e.RData")

# The cleaned ver of Canpath data

@@

#### 1.2 Model mnlm04 same model but for abbott
```{r}
# Model 04  takes sample weights into account
mnlm04<-multinom(relevel(as.factor(time_pso), ref = "never_inf") ~ log_abbott_n, 
                 data = dta_all, weights = sample_wt)
summary(mnlm04)
## use estimates/SE and 1.96 to calculate p-value manually, should be <0.001
```

@@

```{r}
# create by-month estimates, by region
# create by-month est for each region
tb_all<-CBSs %>%
  group_by(samplemonth, region) %>%
  summarise(n = n() , 
            ever_inf= mean(ever_inf, na.rm = T), 
            less_3_month=mean(less_3_month, na.rm = T),
            less_6_month=mean(less_6_month, na.rm=T),
            less_9_month=mean(less_9_month, na.rm=T),
            )
# clean NAs
tb_all<-tb_all[!is.na(tb_all$region), ]
```
#### 2.2.3 Make plots by region

@@

#### 2.3.1 Load & Clean
```{r}
summary(RFD4682_e$N)
# The codebook does not specify what types of anti-N assay is applied, given the value we assume they are using Abbott anti-N assay (small numbers)

# create a subset of APL
APLs<-RFD4682_e %>%
  select(order_ID, AGE_AT_COLLECTION, GENDER, N, `N-IgG_INTERP`, Col_Month, PAT3)

# unify colnames
colnames(APLs)<-c("order_ID","age","gender","abbott_n", "pos_abbott","samplemonth","fsa")
APLs$samplemonth<-as.yearmon(APLs$samplemonth)

# add log_abbott_n
## find 0 values bf transformation
APLs$abbott_n[APLs$abbott_n==0 & 
              !is.na(APLs$abbott_n)]<-0.001

APLs$log_abbott_n <- log(APLs$abbott_n)
```

#### 2.3.2 Apply model (Abbott)
```{r}
# Apply the model mnlm04
APLs_prob_r<-as.data.frame(predict(mnlm04, newdata = APLs, "probs")) # num of >3 months = 0
mean(APLs_prob_r[,1], na.rm = T)  # never: 53.28%
mean(APLs_prob_r[,4], na.rm = T)  # <3 months: 24.21%
mean(APLs_prob_r[,2], na.rm = T)  # 3-6 month: 5.40%
mean(APLs_prob_r[,3], na.rm = T)  # 6-9 months: 4.66%
mean(APLs_prob_r[,5], na.rm = T)  # >9 months: 12.46%

@@

#### 2.3.3 making Plots
```{r}
tb_melt_APL<-reshape2::melt(select(tb_APL, 1,3:6), 
                        id.vars = c("samplemonth"),
                        variable.name = "inf",
                        value.name = "prop")

p06<-tb_melt_APL %>%
  ggplot(aes(x=samplemonth, y=prop, group=inf, colour=inf)) +
  geom_point(aes(colour=inf)) +
  geom_line(aes(colour=inf))  +
  theme_bw()
p06
```
**Additional plotting** 
```{r}
# merge monthly summary table: CBS + APL
tb_CBS<-tb_all[tb_all$region=="Alberta",] %>%
  select(-2)
tb_CBS$data<-rep("CBS", length(tb_CBS$samplemonth))

@@

##===================== Part III. Exp: Reg Model Abbott N~Roche N =============##

This idea came from Alton, we can use the data we have for prob model training to generate another model: Abbott-N~Roche-N. Then apply the model to the CBS data and generate the sero-prevalence curve, see if it gets closer to the APL curve

#### 3.1.1 The linear (log) model: Abbott ~ Roche
```{r}
# The linear model using log_e transformed abbott_n and roche_n

@@

#### 3.1.2  The linear (log) model: Roche ~ Abbott
```{r}
lm01r<-glm(log_roche_n ~ log_abbott_n, 
          family = "gaussian",
          data = dta_all, weights = sample_wt)
summary(lm01r) 
```

#### 3.2.1 Apply the model lm01 to the CBS data
```{r}
log_abbott_n<-predict(lm01, newdata = CBSs, "response")
summary(log_abbott_n)

@@

#### 3.2.2 Apply model lm01r to APL data
```{r}
APLs$log_roche_n<-predict(lm01r, newdata = APLs, "response")

@@

#### 3.3.1 plot: sero-prevalence, using predicted Abbott positivity
```{r}
tb_abb<-CBSs %>%
  group_by(samplemonth, region) %>%
  summarise(n = n() , 
            pos_abbott= sum(pos_abbott, na.rm = T), 
            pos_roche= sum(pos_roche, na.rm = T)
            )
tb_abb$prop.abbott<-tb_abb$pos_abbott/tb_abb$n
tb_abb$prop.roche<-tb_abb$pos_roche/tb_abb$n

@@

#### 3.3.2 plot: sero-prevalence, using predicted Roche pos
```{r}
tb_ro<-APLs %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott= sum(pos_abbott, na.rm = T), 
            pos_roche= sum(pos_roche, na.rm = T)
            )
tb_ro$prop.abbott<-tb_ro$pos_abbott/tb_ro$n
tb_ro$prop.roche<-tb_ro$pos_roche/tb_ro$n

@@

p03r<-tb_ro_melt %>%
  ggplot(aes(x=samplemonth, y=prop, group=assay, colour=assay)) +
  geom_point(aes(colour=assay)) +
  geom_line(aes(colour=assay))  +
  theme_bw() 
p03r
```
#### 3.3.3 Plot both roche and abbott, raw/transformed in on figure
```{r}
# modify melted data

@@

p03.3
```
The gap between CBS and APL seropositivity estimates for October 2023 was X% using raw seropositivity, Y% after transforming CBS to Abbott, and Z% after transforming APL to Roche.

@@

#### 3.4 Plot the correlation btw Roche-N and Abbott-N:
```{r}
# The clear linear relation btw log_roche and log_abbott:
plot(CBSs$log_roche_n, CBSs$log_abbott_n) # this linear association is intriduced by lm01

@@

p04
```
#### Add alternative COI
```{r}
## add two new COI, those two were transformed using the model lm01 and lm01r
df.temp<-as.data.frame(c(log(1.0)))
colnames(df.temp)<-c("log_roche_n")
COI_abbott_t<-predict(lm01, newdata = df.temp, "response")

@@

p05
```
#### Alton suggests using marginal dist with {ggExtra}
```{r}
ggMarginal(p04, groupColour = TRUE, groupFill = TRUE) # log-scale
ggMarginal(p05, groupColour = TRUE, groupFill = TRUE) # with CBS matched data
```
#### 3.5 Create figure panel for AABB conference abstract

#### 3.5.1 Panel A: the serosurveillance figure and RG/Bayesian adjustment
Why: they only alow on figure.
```{r}
## Load Yuan's RG data
rg_all <- read.csv("../1_data/private/df_time_rg_all.csv", row.names=1)
rg_ab  <- rg_all[rg_all$region=="Alberta" &
                 rg_all$source %in% c("APL", "CBS") &
                 !(rg_all$model=="RAW"),]
rg_ab$yearmon <- as.yearmon(rg_ab$yearmon)

@@

#### 3.5.2 Panel B: transformed assay result fig
```{r}
p03.3

@@

#### 3.5.3 Panel C: rolling incidence fig
```{r}
## recorder the dataset, relevel inf variable
tb_melt_2$inf.f<-as.factor(tb_melt_2$inf)
tb_melt_2<-transform(tb_melt_2, 
                     inf.f = factor(inf.f, levels = c("ever_inf", "less_9_month",
                                                     "less_6_month", "less_3_month"))
)
# replace with friendly name for figuers
levels(tb_melt_2$inf.f)<-c("Ever infected", "Infected <9 months ago",
                           "Infected <6 months ago", "Infected <3 months ago")
## make figures p07
p07<-tb_melt_2 %>%
  ggplot(aes(x=samplemonth, y=prop, group=data, colour=data)) +
  geom_point() +
  geom_line()  +
  theme_bw()   +
  xlab("Time of Sample Collection") + 
  ylab("Rolling Incidence Proportion") +
  scale_colour_manual("", values = c("red", "blue"))
p07

@@

#### 3.5.4 Make all three figures into one
```{r}
# use cowplot package
p08.upper<-cowplot::plot_grid(p03.5 + xlab ("") + ylab("Seroprevalence"), 
                              p03.3 + xlab ("") + ylab(""), nrow=1,
                              labels = c("A", "B"))

@@

#### 4.1. About the age dist in training data and CBS: The difference on age distribution might be a potential problem. Generally speaking, blood donors (CBS, FR-Leon data) is relatively older, while Covid-19 convalescent patients are younger.
```{r}
# The FR-Leon data
summary(dta_all[dta_all$study==1,]$age) # mean=46, median = 48
hist(dta_all[dta_all$study==1,]$age)

@@

# KS test: UK+FR-Leon vs CBS
ks.test(dta_all[dta_all$study %in% c(1,6),]$age, 
        CBSs$age[CBSs$age<100])
# as KS test suggests, the distribution of age btw two groups is different
```
```{r}
# create a sampling prob variable, based on the age dist in CBS
CBSs$age_group<-CBS$age_group

@@

#### Create a sample based on the sampling prob:
```{r}
# a simple prob sampling:
set.seed(2048)
dta_sp<-slice_sample(dta_all[!is.na(dta_all$age_w),], n=650,
             weight_by = dta_all[!is.na(dta_all$age_w),]$age_w,
             replace = FALSE)

@@

## Regression on sampled data:
```{r}
mnlm03.s<-multinom(relevel(as.factor(time_pso), ref = "never_inf") ~ log_roche_n + 
                     age + gender, 
                 data = dta_sp,
                 weights = sample_wt)
summary(mnlm03.s)
```

@@

#### 4.2.1 Using Jamaica-Osaka-UK data:
```{r}
# Roche-> Abbott
summary(lm01)

@@

summary(lm01r)
```
#### 4.2.2 Refit using CBS matched data
```{r}
CBSs2<-CBSs %>%
  .[!is.na(.$roche_n) & !is.na(.$abbott_n),] #CBS matched data

@@

# lm05r: Abbott--> Roche, using matched CBS
lm05r<-glm(log_roche_n~log_abbott_n , 
          family = "gaussian",
          data = CBSs2)
summary(lm05r)
# model 5r should be focused
```
#### Adjusting the weights of those with negative test results
```{r}
# add a indicator for positivity of two assays
CBSs2$ppnn[CBSs2$abbott_n > 1.4 & CBSs2$roche_n > 1]<-"pp"
CBSs2$ppnn[CBSs2$abbott_n > 1.4 & CBSs2$roche_n < 1]<-"np"
CBSs2$ppnn[CBSs2$abbott_n <=1.4 & CBSs2$roche_n < 1]<-"nn"
CBSs2$ppnn[CBSs2$abbott_n <=1.4 & CBSs2$roche_n > 1]<-"pn"

@@

>>>>

# How many obs we have in each provinces?
table(CBSs$province, useNA = 'ifany') # n(AB)=124008

# The APL data
load(file = "../1_data/private/RFD4682e.RData")   # ver.1
load(file = "../1_data/private/RFD4682ev2.RData") # ver.2 with additional anti-N data
```

@@

#### 1.2 Model mnlm04 same model but for abbott

```{r}
# Model 04  takes sample weights into account
mnlm04<-multinom(relevel(as.factor(time_pso), ref = "never_inf") ~ log_abbott_n, 
                 data = dta_all, weights = sample_wt)
summary(mnlm04)
## use estimates/SE and 1.96 to calculate p-value manually, should be <0.001
```

@@

```{r}
# create by-month estimates, by region
# create by-month est for each region
tb_all<-CBSs %>%
  group_by(samplemonth, region) %>%
  summarise(n = n() , 
            ever_inf= mean(ever_inf, na.rm = T), 
            less_3_month=mean(less_3_month, na.rm = T),
            less_6_month=mean(less_6_month, na.rm=T),
            less_9_month=mean(less_9_month, na.rm=T),
            )
# clean NAs
tb_all<-tb_all[!is.na(tb_all$region), ]
```

#### 2.2.3 Make plots by region

@@

#### 2.3.1 Load & Clean

```{r}
summary(RFD4682_e$N)
# The codebook does not specify what types of anti-N assay is applied, given the value we assume they are using Abbott anti-N assay (small numbers)

# create a subset of APL
APLs<-RFD4682_ev2 %>%
  select(order_ID, SERO_ID, clean_IDe,
         AGE_AT_COLLECTION, GENDER, `N-IgG_RESULT`,`N-IgG_INTERP`, COLLECTION_DATE)

# unify colnames
colnames(APLs)<-c("order_ID", "SERO_ID", "clean_IDe",
                  "age","gender","abbott_n", "pos_abbott","sampledate")

APLs$samplemonth<-as.yearmon(APLs$sampledate)

# add log_abbott_n
## find 0 values bf transformation, correct them as 0.001
APLs$abbott_n[APLs$abbott_n==0 & 
              !is.na(APLs$abbott_n)]<-0.001

APLs$log_abbott_n <- log(APLs$abbott_n)

# Recode abbott_n test results pos/neg as 1/0
APLs$pos_abbott[APLs$pos_abbott=="Negative"]<-0
APLs$pos_abbott[APLs$pos_abbott=="Positive"]<-1
APLs$pos_abbott<-as.numeric(APLs$pos_abbott)

#### 2.3.2 Apply model (Abbott)

```{r}
# Apply the model mnlm04
APLs_prob_r<-as.data.frame(predict(mnlm04, newdata = APLs, "probs")) # num of >3 months = 0
mean(APLs_prob_r[,1], na.rm = T)  # never: 53.28%
mean(APLs_prob_r[,4], na.rm = T)  # <3 months: 24.21%
mean(APLs_prob_r[,2], na.rm = T)  # 3-6 month: 5.40%
mean(APLs_prob_r[,3], na.rm = T)  # 6-9 months: 4.66%
mean(APLs_prob_r[,5], na.rm = T)  # >9 months: 12.46%

@@

#### 2.3.3 making Plots

```{r}
tb_melt_APL<-reshape2::melt(select(tb_APL, 1,3:6), 
                        id.vars = c("samplemonth"),
                        variable.name = "inf",
                        value.name = "prop")

p06<-tb_melt_APL %>%
  ggplot(aes(x=samplemonth, y=prop, group=inf, colour=inf)) +
  geom_point(aes(colour=inf)) +
  geom_line(aes(colour=inf))  +
  theme_bw()
p06
```

**Additional plotting**

```{r}
# merge monthly summary table: CBS + APL
tb_CBS<-tb_all[tb_all$region=="Alberta",] %>%
  select(-2)
tb_CBS$data<-rep("CBS", length(tb_CBS$samplemonth))

@@

##===================== Part III. Exp: Reg Model Abbott N\~Roche N =============##

This idea came from Alton, we can use the data we have for prob model training to generate another model: Abbott-N\~Roche-N. Then apply the model to the CBS data and generate the sero-prevalence curve, see if it gets closer to the APL curve

#### 3.1.1 The linear (log) model: Abbott \~ Roche

```{r}
# The linear model using log_e transformed abbott_n and roche_n

@@

#### 3.1.2 The linear (log) model: Roche \~ Abbott

```{r}
lm01r<-glm(log_roche_n ~ log_abbott_n, 
          family = "gaussian",
          data = dta_all, weights = sample_wt)
summary(lm01r) 
```

#### 3.2.1 Apply the model lm01 to the CBS data

```{r}
log_abbott_n<-predict(lm01, newdata = CBSs, "response")
summary(log_abbott_n)

@@

#### 3.2.2 Apply model lm01r to APL data

```{r}
APLs$log_roche_n<-predict(lm01r, newdata = APLs, "response")

@@

#### 3.3.1 plot: sero-prevalence, using predicted Abbott positivity

```{r}
tb_abb<-CBSs %>%
  group_by(samplemonth, region) %>%
  summarise(n = n() , 
            pos_abbott= sum(pos_abbott, na.rm = T), 
            pos_roche= sum(pos_roche, na.rm = T)
            )
tb_abb$prop.abbott<-tb_abb$pos_abbott/tb_abb$n
tb_abb$prop.roche<-tb_abb$pos_roche/tb_abb$n

@@

#### 3.3.2 plot: sero-prevalence, using predicted Roche pos

```{r}
tb_ro<-APLs %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott= sum(pos_abbott, na.rm = T), 
            pos_roche= sum(pos_roche, na.rm = T)
            )
tb_ro$prop.abbott<-tb_ro$pos_abbott/tb_ro$n
tb_ro$prop.roche<-tb_ro$pos_roche/tb_ro$n

@@

p03r<-tb_ro_melt %>%
  ggplot(aes(x=samplemonth, y=prop, group=assay, colour=assay)) +
  geom_point(aes(colour=assay)) +
  geom_line(aes(colour=assay))  +
  theme_bw() 
p03r
```

#### 3.3.3 Plot both roche and abbott, raw/transformed in on figure

```{r}
# modify melted data

@@

p03.3
```

The gap between CBS and APL seropositivity estimates for October 2023 was X% using raw seropositivity, Y% after transforming CBS to Abbott, and Z% after transforming APL to Roche.

@@

#### 3.4 Plot the correlation btw Roche-N and Abbott-N:

```{r}
# The clear linear relation btw log_roche and log_abbott:
plot(CBSs$log_roche_n, CBSs$log_abbott_n) # this linear association is intriduced by lm01

@@

p04
```

#### Add alternative COI

```{r}
## add two new COI, those two were transformed using the model lm01 and lm01r
df.temp<-as.data.frame(c(log(1.0)))
colnames(df.temp)<-c("log_roche_n")
COI_abbott_t<-predict(lm01, newdata = df.temp, "response")

@@

p05
```

#### Alton suggests using marginal dist with {ggExtra}

```{r}
ggMarginal(p04, groupColour = TRUE, groupFill = TRUE) # log-scale
ggMarginal(p05, groupColour = TRUE, groupFill = TRUE) # with CBS matched data
```

#### 3.5 Create figure panel for AABB conference abstract

#### 3.5.1 Panel A: the serosurveillance figure and RG/Bayesian adjustment

Why: they only alow on figure.

```{r}
## Load Yuan's RG data
rg_all <- read.csv("../1_data/private/df_time_rg_all.csv", row.names=1)
rg_ab  <- rg_all[rg_all$region=="Alberta" &
                 rg_all$source %in% c("APL", "CBS") &
                 !(rg_all$model=="RAW"),]
rg_ab$yearmon <- as.yearmon(rg_ab$yearmon)

@@

#### 3.5.2 Panel B: transformed assay result fig

```{r}
p03.3

@@

#### 3.5.3 Panel C: rolling incidence fig

```{r}
## recorder the dataset, relevel inf variable
tb_melt_2$inf.f<-as.factor(tb_melt_2$inf)
tb_melt_2<-transform(tb_melt_2, 
                     inf.f = factor(inf.f, levels = c("ever_inf", "less_9_month",
                                                     "less_6_month", "less_3_month"))
)
# replace with friendly name for figuers
levels(tb_melt_2$inf.f)<-c("Ever infected", "Infected <9 months ago",
                           "Infected <6 months ago", "Infected <3 months ago")
## make figures p07
p07<-tb_melt_2 %>%
  ggplot(aes(x=samplemonth, y=prop, group=data, colour=data)) +
  geom_point() +
  geom_line()  +
  theme_bw()   +
  xlab("Time of Sample Collection") + 
  ylab("Rolling Incidence Proportion") +
  scale_colour_manual("", values = c("red", "blue"))
p07

@@

#### 3.5.4 Make all three figures into one

```{r}
# use cowplot package
p08.upper<-cowplot::plot_grid(p03.5 + xlab ("") + ylab("Seroprevalence"), 
                              p03.3 + xlab ("") + ylab(""), nrow=1,
                              labels = c("A", "B"))

@@

#### 4.1. About the age dist in training data and CBS: The difference on age distribution might be a potential problem. Generally speaking, blood donors (CBS, FR-Leon data) is relatively older, while Covid-19 convalescent patients are younger.

```{r}
# The FR-Leon data
summary(dta_all[dta_all$study==1,]$age) # mean=46, median = 48
hist(dta_all[dta_all$study==1,]$age)

@@

# KS test: UK+FR-Leon vs CBS
ks.test(dta_all[dta_all$study %in% c(1,6),]$age, 
        CBSs$age[CBSs$age<100])
# as KS test suggests, the distribution of age btw two groups is different
```

```{r}
# create a sampling prob variable, based on the age dist in CBS
CBSs$age_group<-CBS$age_group

@@

#### Create a sample based on the sampling prob:

```{r}
# a simple prob sampling:
set.seed(2048)
dta_sp<-slice_sample(dta_all[!is.na(dta_all$age_w),], n=650,
             weight_by = dta_all[!is.na(dta_all$age_w),]$age_w,
             replace = FALSE)

@@

## Regression on sampled data:

```{r}
mnlm03.s<-multinom(relevel(as.factor(time_pso), ref = "never_inf") ~ log_roche_n + 
                     age + gender, 
                 data = dta_sp,
                 weights = sample_wt)
summary(mnlm03.s)
```

@@

#### 4.2.1 Using Jamaica-Osaka-UK data:

```{r}
# Roche-> Abbott
summary(lm01)

@@

summary(lm01r)
```

#### 4.2.2 Refit using CBS matched data

```{r}
CBSs2<-CBSs %>%
  .[!is.na(.$roche_n) & !is.na(.$abbott_n),] #CBS matched data

@@

# lm05r: Abbott--> Roche, using matched CBS
lm05r<-glm(log_roche_n~log_abbott_n , 
          family = "gaussian",
          data = CBSs2)
summary(lm05r)
# model 5r should be focused
```

#### Adjusting the weights of those with negative test results

```{r}
# add a indicator for positivity of two assays
CBSs2$ppnn[CBSs2$abbott_n > 1.4 & CBSs2$roche_n > 1]<-"pp"
CBSs2$ppnn[CBSs2$abbott_n > 1.4 & CBSs2$roche_n < 1]<-"np"
CBSs2$ppnn[CBSs2$abbott_n <=1.4 & CBSs2$roche_n < 1]<-"nn"
CBSs2$ppnn[CBSs2$abbott_n <=1.4 & CBSs2$roche_n > 1]<-"pn"

@@

#### 4.2.3 Add regression lines to the figure

```{r}
# Modify p04 for sensitivity analysis
p04<-dta_all %>%
  .[!is.na(.$log_abbott_n) & !is.na(.$log_roche_n),] %>%
  ggplot(aes(x=log_roche_n, y=log_abbott_n, 
             study, colour=study)) +
  geom_point(aes(colour=as.factor(study)), alpha=0.5) +
  geom_vline(xintercept = 0, linewidth=1.0, linetype="dotted", col="darkgrey") +
  geom_hline(yintercept = log(1.4), linewidth=1.0, linetype="dotted", col="darkgrey") +
  xlab("Roche Anti-N S/Co ratio (log scale)") + 
  ylab("Abbott Anti-N S/Co ratio (log scale)") +
  labs(color='') +
  scale_colour_manual(labels = c("Jamaica (-)", "Osaka, JP (-)",
                                 "Osaka, JP (+)","UK (+)"), 
                      values = c("red", "blue", "darkgreen", "purple")) +
  theme_bw() + 
  theme(legend.position = "bottom")

p04.2a<-p04 + geom_abline(aes(intercept = -1.60, slope = 0.68, color="Roche->Abbott",),
                  size=0.8) +
      geom_abline(aes(intercept = -1.68, slope = 0.78, color="Abbott->Roche",), 
                  size=0.8) +
scale_colour_manual(values = c("red", "blue", "darkgreen", "purple","blue","red"),
                    guide = guide_legend(nrow = 3,
                            override.aes = list(
                            linetype = c(rep("blank", 4), "solid", "solid"),
                            shape = c(rep(16, 4), NA, NA)))) +
     ggtitle("Jamaica-Osaka-UK matched data")
```

```{r}
# Calculating Beta estimates for model 01r-reverted
(-1)*lm01r$coefficients[1]/lm01r$coefficients[2] #beta_0
1/lm01r$coefficients[2]                          #beta_1

# Calculating Beta estimates for model 05r-reverted
(-1)*lm05r$coefficients[1]/lm05r$coefficients[2] #beta_0
1/lm05r$coefficients[2]                          #beta_1

# Calculating Beta estimates for model 05rw-reverted
(-1)*lm05r.w$coefficients[1]/lm05r.w$coefficients[2] #beta_0
1/lm05r.w$coefficients[2]                            #beta_1
```

```{r}
# Refit models for CBS data, modify p05
p04.2b<-dta_CBS %>%
  .[!is.na(.$log_abbott_n) & !is.na(.$log_roche_n) &
    .$study=="CBS matched",] %>%
  ggplot(aes(x=log_roche_n, y=log_abbott_n, 
             study, colour=study)) +
  geom_point(aes(colour=as.factor(study)), alpha=0.40) +
  geom_vline(xintercept = 0, linewidth=1.0, linetype="dotted", col="darkgrey") +
  geom_hline(yintercept = log(1.4), linewidth=1.0, linetype="dotted", col="darkgrey") +
  xlab("Roche Anti-N S/Co ratio (log scale)") + 
  ylab("Abbott Anti-N S/Co ratio (log scale)") +
  labs(color='') +
  scale_colour_manual(labels = c("CBS matched"), 
                      values = c("lightblue")) +
  theme_bw() + 
  theme(legend.position = "bottom")

p04.2b<-p04.2b + geom_abline(aes(intercept = -1.87, slope = 0.64, color="Roche->Abbott",),
                  size=0.8) +
         geom_abline(aes(intercept = 2.02, slope = 2.26, color="Abbott->Roche",), 
                  size=0.8) +
         geom_abline(aes(intercept = -2.14, slope = 0.85, 
                         color="Abbott->Roche (down-weighted)",), 
                  size=0.8, linetype=2) +
         scale_colour_manual(values = c("lightblue","blue","red","red"),
                             guide = guide_legend(nrow=2,
                             override.aes = list(
                             linetype = c(rep("blank", 1), "solid", "solid","dashed"),
                             shape = c(rep(16, 1), NA, NA, NA))
                             )) +
         ggtitle("CBS matched data")
p04.2b
```

```{r}
## make two into one paneled fig
png(file="../4_output/figs/Fig.S1 Sensitivity Analysis.png", 
    width=2400, height=1440, res=300)
cowplot::plot_grid(p04.2a, 
                   p04.2b  + ylab(""), nrow=1,
                   align = 'h',
                   labels = c("A", "B"))
dev.off()
```

#### 4.3 RG Adjusted Sero-prevalence using Bayesian model's priors

##### 4.3.1 Create monthly anti-N positivity data for CBS
```{r}
# CBSs AB subset, N=124008
CBSs_AB<-CBSs[CBSs$province == "AB",]

# Monthly Positivity in CBS
tb_431<-CBSs_AB %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_roche= sum(pos_roche, na.rm = T)
            )
tb_431$prop.raw<-tb_431$pos_roche/tb_431$n

# Add RG adjustment
tb_431$prop.adj.p1<-(tb_431$prop.raw + 0.998 -1)/(0.998 + 0.979 -1)
tb_431$prop.adj.p2<-(tb_431$prop.raw + 0.993 -1)/(0.993 + 0.862 -1)
tb_431$study <- rep("CBS", length(tb_431$samplemonth))
```

##### 4.3.2 Create monthly anti-N positivity data for APL
```{r}
# APL data, N=214776

# Monthly Positivity in APL
tb_432<-APLs %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott= sum(pos_abbott, na.rm = T)
            )
tb_432$prop.raw<-tb_432$pos_abbott/tb_432$n

# remove missing data
tb_432<-tb_432[!is.na(tb_432$samplemonth),]
tb_432<-tb_432[tb_432$pos_abbott!=0,]

# Add RG adjustment
tb_432$prop.adj.p1<-(tb_432$prop.raw + 0.995 -1)/(0.995 + 0.960 -1)
tb_432$prop.adj.p2<-(tb_432$prop.raw + 0.995 -1)/(0.995 + 0.838 -1)
tb_432$study <- rep("APL", length(tb_432$samplemonth))

# merge CBS and APL table
tb_43<-bind_rows(tb_431[, c(1, 4:7)], tb_432[-c(1), c(1, 4:7)])
# April 2020 row is removed

tb_43m<-reshape2::melt(tb_43, 
                        id.vars = c("samplemonth", "study"),
                        variable.name = "adjustment",
                        value.name = "prop")
rm(tb_431, tb_432)
```

##### 4.3.3 Make plots
```{r}
p4.3.3<-tb_43m %>%
  ggplot(aes(x=samplemonth, y=prop, colour=study)) +
  geom_point(aes(colour=study)) +
  geom_line(aes(colour=study, linetype=adjustment))  +
  xlab("Month of Sample Collection") +
  ylab("Seroprevalence") +
  coord_cartesian(xlim = c(as.yearmon("May 2020"), as.yearmon("Jan 2022")),
                  ylim = c(0, 0.15)) +
#  ylim(c(as.yearmon("Jan 2022"), as.yearmon("Feb 2023"))) +
  scale_colour_manual("", values = c("red","blue")) +
  scale_linetype_manual("", values=c(1,2,3),
                        labels= c("Unadjusted", "RG adjusted (informed priors)",
                                  "RG adjusted (nautral priors)")) +
  theme_bw()

p4.3.3
```
